Spatial Effort Analaysis
========================================================
###2013-9-18

Import data from arcGIS:

```{r}
#####
#Clear the junk
graphics.off()
rm(list=ls(all=TRUE))

#setwd("C:/Users/kevin.purcell/Documents/NOAA_SpatialEffort_MS/mapping_2013-3-5")

library(reshape)
library(mgcv)  #loading the Woods GAM package
library(lattice)
library(maptools)
library(maps)
library(mapdata)

#import new dbf file from spatial join off polygon 2 points
#this data attaches centroid locations for 10x10 min grid cells with all 
#6-8 elb data falling within those polygons

library(foreign)

elb.2mo.dat<-read.dbf(file="C:/Users/kevin.purcell/Documents/NOAA_SpatialEffort_MS/mapping_2013-3-5/poly2point_2mo_sz.dbf", as.is = FALSE)

summary(elb.2mo.dat)

elb.2mo.dat<-subset(elb.2mo.dat, elb.2mo.dat$depth<=100)
```

This dataset was preprocessed in GIS by:
* Importing the raw ELB data and spatially mapping fishing locations
* Overlaying using the fishnet function in arcCatalog a 10 x 10 min grid
* These two layers were joined by assinging each fishing location to a grid cell
* SEAMAP survey environmental and depth data were imported into arcMAP
* polygon fishing effort was match with environmental data at 2 mon resolution
* ADD MORE METHODS



The effort of the fishery is a factor in the model.  We have two sources of data for fishery effort the ELB and the SHRCOM data.

```{r}
tot.eff<-aggregate(towhours ~ yr, data=elb.2mo.dat, FUN="sum") #ELB eff agg per JCK discuss
colnames(tot.eff)[2]<-'tot.eff'

plot(tot.eff$tot.eff~ tot.eff$yr, type="b",
     ylab="Shrimping Effort (tow hours)",
     xlab="Year",
     main="ELB Effort Time Series")
grid(col="gray")

```

ELB data on effort shows an increase from 2004-2009 with a big decline in 2010 and then a subsequent increase in 2010.  My thinking is that this represents sampling rather than the behavior of the fishery.  



```{r}
# import shrcom time series and examine
shrcom.dat<-read.csv("C:/Users/kevin.purcell/Documents/comm_analysis/offshore_6011.csv", sep=",",header=T)
shrcom.dat$hrs<-(shrcom.dat$effort*24)
shrcom.sum<-subset(shrcom.dat, shrcom.dat$Season==2)
shrcom.sum<-subset(shrcom.dat, shrcom.dat$area>=3)
shrcom.yr<-aggregate(hrs ~ yr, data=shrcom.sum, FUN="sum")
shrcom.yr<-subset(shrcom.yr, shrcom.yr$yr>=2004&shrcom.yr$yr<=2010)

plot(shrcom.yr$hrs~ shrcom.yr$yr, type="b",
     ylab="Shrimping Effort (tow hours)",
     xlab="Year",
     main="SHRCOM Effort Time Series")
grid(col="gray")
names(shrcom.yr)
shrcom.mean<-mean(shrcom.yr$hrs)
shrcom.yr$hrs<-(shrcom.yr$hrs/shrcom.mean)
```

The SHRCOM data shows sharp decline from 2004-2008 at which point it seems to stabalize. It seems that this time series is more representative of the fishery.  


```{r}

# shrimp abundance data itegrated to the 10x10 grid.
abun.dat<-read.csv("C:/Users/kevin.purcell/Documents/NOAA_SpatialEffort_MS/grid2cpue.csv", sep=",",header=T)

abun.dat<-subset(abun.dat, abun.dat$month>=6&abun.dat$month<=8)
summary(abun.dat)
#create a calendar week vector
abun.dat$DATE<-paste(abun.dat$YR, abun.dat$month, sep="-")
abun.dat$DATE<-paste(abun.dat$DATE, abun.dat$day, sep="-")
#abun.dat$DATE <- as.POSIXlt(abun.dat$DATE)
abun.dat$CW <- strftime(abun.dat$DATE,format="%W")
abun.dat$jd <- strftime(abun.dat$DATE, format="%j")
abun.dat$CW<-as.numeric(abun.dat$CW)
abun.dat$jd<-as.numeric(abun.dat$jd)

abun.avg=aggregate(cpue ~ cent_lon + cent_lat + YR,
                       data=abun.dat, FUN=mean)
summary(abun.avg)
#boxplot(abun.avg$cpue ~ abun.avg$YR)
#hist(abun.avg$cpue)
#stem(abun.avg$cpue)
#abun.avg<-subset(abun.avg, abun.avg$cpue<10)


```

```{r}
#fuel price data
fuel.dat<-read.table("C:/Users/kevin.purcell/Documents/NOAA_SpatialEffort_MS/la_fuel_price.csv", sep=",", header=T)
fuel.dat$DATE<-paste(fuel.dat$yr, fuel.dat$mo, sep="-")
fuel.dat$DATE<-paste(fuel.dat$DATE, fuel.dat$day, sep="-")
fuel.dat$CW <- strftime(fuel.dat$DATE,format="%W")
fuel.dat$jd <- strftime(fuel.dat$DATE, format="%j")
fuel.dat$CW<-as.numeric(fuel.dat$CW)
fuel.dat$jd<-as.numeric(fuel.dat$jd)
fuel.agg<-aggregate(la_fuel_price ~ yr, data=fuel.dat, FUN=mean)

boxplot(fuel.dat$la_fuel_price~fuel.dat$yr,
        main="Fuel Price Variation")
grid(col="gray")

#plot
plot(fuel.agg$la_fuel_price~fuel.agg$yr, type="b",
     ylab="Fuel Price",
     xlab="Year",
     main="Fuel Price Timeseries for Los Angeles")
grid(col="gray")
```

```{r}
#price data
ppnd.dat=read.csv("C:/Users/kevin.purcell/Documents/NOAA_SpatialEffort_MS/ppnd.csv", sep=",",header=T)
summary(ppnd.dat)
ppnd.dat$PRICEPPND<-ppnd.dat$PRICEPPND/100
colnames(ppnd.dat)[1]<-'sz'
colnames(ppnd.dat)[2]<-'yr'
colnames(ppnd.dat)[5]<-'ppnd'
#limit ppnd data
ppnd.dat<-subset(ppnd.dat, ppnd.dat$ppnd<=6)
ppnd.yr<-aggregate(ppnd~yr, data=ppnd.dat, FUN=mean)


boxplot(ppnd.dat$ppnd~ppnd.dat$yr,
        main="Shrimp Price Variation")
grid(col="gray")


plot(ppnd.yr$ppnd~ppnd.yr$yr, type="b",
     ylab="Price per Pond (USD)",
     xlab="Year",
     main="Shrimp Price Time Series")
grid(col="gray")



#remove data points where centroid data is missing
elb.2mo.dat<- subset(elb.2mo.dat, elb.2mo.dat$cent_lon!=0)
elb.2mo.dat<- subset(elb.2mo.dat, elb.2mo.dat$ZONE!=0)
#elb.2mo.dat = 149840

#rename some columns in the new dataset
colnames(elb.2mo.dat)[12]<-'grid_cell'
colnames(elb.2mo.dat)[3]<-'vesselID'
colnames(elb.2mo.dat)[10]<-'depth'
colnames(elb.2mo.dat)[17]<-'sz'
elb.2mo.dat$FID_1<-NULL
elb.2mo.dat$FID_3<-NULL
elb.2mo.dat$FID_1_1<-NULL
elb.2mo.dat$Id<-NULL
                 
#create date stuff
#create a calendar week vector
elb.2mo.dat$DATE<-paste(elb.2mo.dat$yr, elb.2mo.dat$mon, sep="-")
elb.2mo.dat$DATE<-paste(elb.2mo.dat$DATE, elb.2mo.dat$day, sep="-")
#elb.2mo.dat$DATE <- as.POSIXlt(elb.2mo.dat$DATE)
elb.2mo.dat$CW <- strftime(elb.2mo.dat$DATE,format="%W")
elb.2mo.dat$jd <- strftime(elb.2mo.dat$DATE, format="%j")
elb.2mo.dat$CW<-as.numeric(elb.2mo.dat$CW)
elb.2mo.dat$jd<-as.numeric(elb.2mo.dat$jd)
summary(elb.2mo.dat)

```


```{r}
hypox.area=read.table("C:/Users/kevin.purcell/Documents/NOAA_SpatialEffort_MS/analysis/area_of_hypoxia_NR.csv", sep=",",header=T)
colnames(hypox.area)[1]<-'yr'

plot(hypox.area$area.hypoxic~hypox.area$yr, type="b",
     ylab="Areial Extent of Hypoxia (sq km)",
     xlab="Year",
     main="Rabalais Hypoxia Time Series")
grid(col="gray")

```


Average out the environmental data per cell and integrate the data sets into modeling data sets

```{r}

summary(elb.2mo.dat)

#aggregate response variables to polygons

grid.2mo.avg=aggregate(cbind(do, depth) ~ grid_cell * cent_lon * cent_lat* yr * jd,
                       data=elb.2mo.dat, FUN=mean)
tx.cells=length(grid.2mo.avg$do[grid.2mo.avg$cent_lon<(-94)])
la.cells=length(grid.2mo.avg$do[grid.2mo.avg$cent_lon>(-94)])



#aggregate 2mo responses to grid cells by jd
g2mo.dur<-aggregate(towhours ~ grid_cell * cent_lon * cent_lat * yr * jd ,
                    data=elb.2mo.dat, FUN=sum)
g2mo.avg.dur<-aggregate(towhours ~ grid_cell * cent_lon * cent_lat * yr * jd,
                        data=elb.2mo.dat, FUN=mean)
colnames(g2mo.avg.dur)[6]<-'avg.dur'
g2mo.tow.count<-aggregate(towhours ~ grid_cell * cent_lon * cent_lat* yr * jd,
                          data=elb.2mo.dat, FUN=length)
colnames(g2mo.tow.count)[6]<-'tow.cnt'
g2mo.vess.count<-aggregate(vesselID ~ grid_cell * cent_lon * cent_lat* yr * jd,
                           data=elb.2mo.dat, FUN=length)
colnames(g2mo.vess.count)[6]<-'vess.cnt'



#merge with avg environmental conditions
gam.2mo.dur<-merge(g2mo.dur, grid.2mo.avg, all.x=T,
                   by=c("grid_cell", "cent_lon", "cent_lat", "yr", "jd"))
colnames(abun.avg)[3]<-'yr'
gam.2mo.dur<-merge(gam.2mo.dur, abun.avg, all.x=T,
                   by=c("cent_lon", "cent_lat", "yr"))


gam.2mo.dur$sz<-gam.2mo.dur$grid_cell
gam.2mo.dur$sz[gam.2mo.dur$cent_lon>-90&gam.2mo.dur$cent_lon<=-89]=13
gam.2mo.dur$sz[gam.2mo.dur$cent_lon>-91&gam.2mo.dur$cent_lon<=-90]=14
gam.2mo.dur$sz[gam.2mo.dur$cent_lon>-92&gam.2mo.dur$cent_lon<=-91]=15
gam.2mo.dur$sz[gam.2mo.dur$cent_lon>-93&gam.2mo.dur$cent_lon<=-92]=16
gam.2mo.dur$sz[gam.2mo.dur$cent_lon>-94&gam.2mo.dur$cent_lon<=-93]=17

gam.2mo.dur$sz[gam.2mo.dur$cent_lon>-95&gam.2mo.dur$cent_lon<=-94]=18
gam.2mo.dur$sz[gam.2mo.dur$cent_lon<=-95&gam.2mo.dur$cent_lat>28]=19
gam.2mo.dur$sz[gam.2mo.dur$cent_lat>=27&gam.2mo.dur$cent_lat<28]=20
gam.2mo.dur$sz[gam.2mo.dur$cent_lat>=26&gam.2mo.dur$cent_lat<27]=21

gam.2mo.dur<-merge(gam.2mo.dur, ppnd.dat,all.x=T,
                   by=c("yr", "jd", "sz"))

gam.2mo.dur<-merge(gam.2mo.dur, hypox.area,all.x=T,
                   by="yr")
gam.2mo.dur<-merge(gam.2mo.dur, shrcom.yr, all.x=T,
                       by="yr")

fuel.agg<-aggregate(la_fuel_price ~ CW + yr, data=fuel.dat, FUN=mean)
gam.2mo.dur=merge(gam.2mo.dur, fuel.agg, by=c("CW","yr"), all.x=T)

summary(gam.2mo.dur)

#merge 
gam.2mo.avg.dur<-merge(g2mo.avg.dur, grid.2mo.avg, all.x=T,
                   by=c("grid_cell", "cent_lon", "cent_lat", "yr", "jd"))
colnames(abun.avg)[3]<-'yr'
gam.2mo.avg.dur<-merge(gam.2mo.avg.dur, abun.avg, all.x=T,
                   by=c("cent_lon", "cent_lat", "yr"))

gam.2mo.avg.dur$sz<-gam.2mo.avg.dur$grid_cell
gam.2mo.avg.dur$sz[gam.2mo.avg.dur$cent_lon>-90&gam.2mo.avg.dur$cent_lon<=-89]=13
gam.2mo.avg.dur$sz[gam.2mo.avg.dur$cent_lon>-91&gam.2mo.avg.dur$cent_lon<=-90]=14
gam.2mo.avg.dur$sz[gam.2mo.avg.dur$cent_lon>-92&gam.2mo.avg.dur$cent_lon<=-91]=15
gam.2mo.avg.dur$sz[gam.2mo.avg.dur$cent_lon>-93&gam.2mo.avg.dur$cent_lon<=-92]=16
gam.2mo.avg.dur$sz[gam.2mo.avg.dur$cent_lon>-94&gam.2mo.avg.dur$cent_lon<=-93]=17

gam.2mo.avg.dur$sz[gam.2mo.avg.dur$cent_lon>-95&gam.2mo.avg.dur$cent_lon<=-94]=18
gam.2mo.avg.dur$sz[gam.2mo.avg.dur$cent_lon<=-95&gam.2mo.avg.dur$cent_lat>28]=19
gam.2mo.avg.dur$sz[gam.2mo.avg.dur$cent_lat>=27&gam.2mo.avg.dur$cent_lat<28]=20
gam.2mo.avg.dur$sz[gam.2mo.avg.dur$cent_lat>=26&gam.2mo.avg.dur$cent_lat<27]=21

gam.2mo.avg.dur<-merge(gam.2mo.avg.dur, ppnd.dat,all.x=T,
                   by=c("yr", "jd", "sz"))

gam.2mo.avg.dur<-merge(gam.2mo.avg.dur, hypox.area,all.x=T,
                   by="yr")
gam.2mo.avg.dur<-merge(gam.2mo.avg.dur, shrcom.yr, all.x=T,
                          by="yr")
gam.2mo.avg.dur=merge(gam.2mo.avg.dur, fuel.agg, by="CW", all.x=T)

#merge with avg environmental conditions
gam.2mo.tow.count<-merge(g2mo.tow.count, grid.2mo.avg, all.x=T,
                   by=c("grid_cell", "cent_lon", "cent_lat", "yr", "jd"))
colnames(abun.avg)[3]<-'yr'
gam.2mo.tow.count<-merge(gam.2mo.tow.count, abun.avg, all.x=T,
                   by=c("cent_lon", "cent_lat", "yr"))

gam.2mo.tow.count$sz<-gam.2mo.tow.count$grid_cell
gam.2mo.tow.count$sz[gam.2mo.tow.count$cent_lon>-90&gam.2mo.tow.count$cent_lon<=-89]=13
gam.2mo.tow.count$sz[gam.2mo.tow.count$cent_lon>-91&gam.2mo.tow.count$cent_lon<=-90]=14
gam.2mo.tow.count$sz[gam.2mo.tow.count$cent_lon>-92&gam.2mo.tow.count$cent_lon<=-91]=15
gam.2mo.tow.count$sz[gam.2mo.tow.count$cent_lon>-93&gam.2mo.tow.count$cent_lon<=-92]=16
gam.2mo.tow.count$sz[gam.2mo.tow.count$cent_lon>-94&gam.2mo.tow.count$cent_lon<=-93]=17

gam.2mo.tow.count$sz[gam.2mo.tow.count$cent_lon>-95&gam.2mo.tow.count$cent_lon<=-94]=18
gam.2mo.tow.count$sz[gam.2mo.tow.count$cent_lon<=-95&gam.2mo.tow.count$cent_lat>28]=19
gam.2mo.tow.count$sz[gam.2mo.tow.count$cent_lat>=27&gam.2mo.tow.count$cent_lat<28]=20
gam.2mo.tow.count$sz[gam.2mo.tow.count$cent_lat>=26&gam.2mo.tow.count$cent_lat<27]=21

gam.2mo.tow.count<-merge(gam.2mo.tow.count, ppnd.dat,all.x=T,
                   by=c("yr", "jd", "sz"))

gam.2mo.tow.count<-merge(gam.2mo.tow.count, hypox.area,all.x=T,
                   by="yr")
gam.2mo.tow.count<-merge(gam.2mo.tow.count, shrcom.yr, all.x=T,
                          by="yr")
names(gam.2mo.tow.count)
colnames(gam.2mo.tow.count)[6]<-'CW'
colnames(gam.2mo.tow.count)[11]<-'ppnd'
colnames(gam.2mo.tow.count)[12]<-'area.hypoxic'
colnames(gam.2mo.tow.count)[13]<-'hrs'
gam.2mo.tow.count=merge(gam.2mo.tow.count, fuel.agg, by="CW", all.x=T)


#make two data sets a LA and TX
gam.2mo.dur.la<-subset(gam.2mo.dur, gam.2mo.dur$cent_lon>=-93.83)
gam.2mo.dur.tx<-subset(gam.2mo.dur, gam.2mo.dur$cent_lon< -93.83)
gam.2mo.avg.dur.la<-subset(gam.2mo.avg.dur, gam.2mo.avg.dur$cent_lon>=-93.83)
gam.2mo.avg.dur.tx<-subset(gam.2mo.avg.dur, gam.2mo.avg.dur$cent_lon< -93.83)
gam.2mo.tow.count.la<-subset(gam.2mo.tow.count, gam.2mo.tow.count$cent_lon>=-93.83)
gam.2mo.tow.count.tx<-subset(gam.2mo.tow.count, gam.2mo.tow.count$cent_lon< -93.83)
# gam.2mo.vess.count.la<-subset(gam.2mo.vess.count, gam.2mo.vess.count$cent_lon>=-93.83)
# gam.2mo.vess.count.tx<-subset(gam.2mo.vess.count, gam.2mo.vess.count$cent_lon< -93.83)

save.image("C:/Users/kevin.purcell/Documents/NOAA_SpatialEffort_MS/GAM_2013-9-19.RData")



#total effort or duration in a cell calculated by FUN=sum
summary(gam.2mo.dur)

#make region code
gam.2mo.dur$region<-gam.2mo.dur$sz
gam.2mo.dur$region[gam.2mo.dur$sz>=13&gam.2mo.dur$sz<=17]="LA"
gam.2mo.dur$region[gam.2mo.dur$sz>=18&gam.2mo.dur$sz<=21]="TX"                 
 
gam.dur<-gam.2mo.dur[,c("yr","jd","hrs","do","depth", "ppnd", "la_fuel_price","towhours","area.hypoxic","cent_lon","cent_lat")]    

summary(gam.dur)
par(mfrow=c(2,2))
hist(gam.dur$yr)
hist(gam.dur$jd)
boxplot(gam.dur$do~gam.dur$yr)
boxplot(gam.dur$depth~gam.dur$yr)

gam.dur.la<-gam.2mo.dur.la[,c("jd", "hrs","do","depth", "ppnd", "la_fuel_price","towhours","area.hypoxic","cent_lon", "cent_lat")]
gam.dur.la<-na.omit(gam.dur.la)
# 
# gam.dur.tx<-gam.2mo.dur.tx[,c("jd", "hrs","do","depth", "ppnd", "la_fuel_price","towhours","area.hypoxic","cent_lon", "cent_lat")]
# gam.dur.tx<-na.omit(gam.dur.tx)
```

```{r}
#remove na's
gam.dur<-na.omit(gam.dur)
gam.dur.2004<-subset(gam.dur, gam.dur$yr==2004)
gam.dur.2005<-subset(gam.dur, gam.dur$yr==2005)
gam.dur.2006<-subset(gam.dur, gam.dur$yr==2006)
gam.dur.2007<-subset(gam.dur, gam.dur$yr==2007)
gam.dur.2008<-subset(gam.dur, gam.dur$yr==2008)
gam.dur.2009<-subset(gam.dur, gam.dur$yr==2009)
gam.dur.2010<-subset(gam.dur, gam.dur$yr==2010)
gam.dur.2010<-subset(gam.dur, gam.dur$yr==2010)


#GOM model
gom1<-gam(log(towhours+1) ~ factor(yr) + la_fuel_price  + hrs + area.hypoxic + 
          s(do) + s(depth) + s(ppnd) + s(jd) + s(cent_lon,cent_lat), 
          data=gam.dur)

gom2<-gam(log(towhours+1) ~ factor(yr) + la_fuel_price  + hrs + area.hypoxic + 
          s(do) + s(depth) + s(ppnd) + s(jd) + s(cent_lon,cent_lat) + 
          s(cent_lon,cent_lat, by=area.hypoxic) + s(cent_lon,cent_lat, by=hrs),
          data=gam.dur)

#
gam.check(gom1)
gam.check(gom2)


summary(gom1)
summary(gom2)

par(mfrow=c(2,2))
plot(gom2, select=1, shade=T, all.terms=T, scale=0,
     xlab="Dissolved Oxygen (mg/L)")
text(.4, -0.72, "A", cex=1.5,font=2)
plot(gom2, select=2, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
text(3, -2, "B", cex=1.5,font=2)
plot(gom2, select=3, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
text(0.4, -0.6, "C", cex=1.5,font=2)
plot(gom2, select=4, shade=T, all.terms=T, scale=0,
     xlab="Julian Day" )
text(1.563, -0.44, "D", cex=1.5,font=2)

plot(gom2, select=6, se=F, rug=F, scheme=2,
     xlab="Longitude",
     ylab="Latitude",
     main="Effect of area of hypoxia") 
map("state",fill=T,col="gray",add=T)

plot(gom2, select=7, se=F, rug=F, scheme=2,
     xlab="Longitude",
     ylab="Latitude",
     main="Effect of effort") 
map("state",fill=T,col="gray",add=T)



#Get slope coefficients for gom2 model
pred<-predict(gom2, type='terms')
pred.slope<-(pred[,10]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))

par(mfrow=c(1,2))
vis.gam(gom2, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur$cent_lon, gam.dur$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur$cent_lon[pred.slope>=0],gam.dur$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='red')
symbols(gam.dur$cent_lon[pred.slope<=0],gam.dur$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='blue')
map("worldHires", fill=T, col="grey",add=T)
```


```{r}
#old model
old1<-gam(log(towhours+1) ~ s(jd) + s(depth) + s(ppnd) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=area.hypoxic),data=gam.dur)
old2<-gam(log(towhours+1) ~ jd + hrs + s(do) + s(depth) + s(ppnd) + s(la_fuel_price) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=area.hypoxic) + s(cent_lon,cent_lat, by=hrs),
         data=gam.dur)

summary(old1)
summary(old2)


par(mfrow=c(2,3))

plot(old1, select=2, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
text(3, -2, "B", cex=1.5,font=2)
plot(old1, select=3, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
text(0.4, -0.6, "C", cex=1.5,font=2)
plot(old1, select=4, shade=T, all.terms=T, scale=0,
     xlab="Fuel Price (USD)" )
text(1.563, -0.44, "D", cex=1.5,font=2)
plot(old1, select=1, shade=T, all.terms=T, scale=0,
     xlab="Julian Day")
text(.4, -0.72, "A", cex=1.5,font=2)

par(mfrow=c(2,2))
plot(old2, select=1, shade=T, all.terms=T, scale=0,
     xlab="Dissolved Oxygen (mg/L)")
text(.4, -0.72, "A", cex=1.5,font=2)
plot(old2, select=2, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
text(3, -2, "B", cex=1.5,font=2)
plot(old2, select=3, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
text(0.4, -0.6, "C", cex=1.5,font=2)
plot(old2, select=4, shade=T, all.terms=T, scale=0,
     xlab="Fuel Price (USD)" )
text(1.563, -0.44, "D", cex=1.5,font=2)


#Get slope coefficients for gom2 model
pred<-predict(old1, type='terms')
head(pred)
pred.slope<-(pred[,7]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))
par(mfrow=c(1,2))
vis.gam(old1, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur$cent_lon, gam.dur$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur$cent_lon[pred.slope>=0],gam.dur$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='blue')
symbols(gam.dur$cent_lon[pred.slope<=0],gam.dur$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='red')
map("worldHires", fill=T, col="grey",add=T)
```

```{r}
#gams by year

m2004<-gam(log(towhours+1) ~ area.hypoxic + s(jd) +  s(depth) + s(ppnd) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=do),data=gam.dur.2004)
m2005<-gam(log(towhours+1) ~ area.hypoxic +s(jd) + s(depth) + s(ppnd) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=do),data=gam.dur.2005)
m2006<-gam(log(towhours+1) ~ area.hypoxic +s(jd) + s(depth) + s(ppnd) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=do),data=gam.dur.2006)
m2007<-gam(log(towhours+1) ~ area.hypoxic +s(jd) + s(depth) + s(ppnd) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=do),data=gam.dur.2007)
m2008<-gam(log(towhours+1) ~ area.hypoxic +s(jd) + s(depth) + s(ppnd) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=do),data=gam.dur.2008)
m2009<-gam(log(towhours+1) ~ area.hypoxic +s(jd) + s(depth) + s(ppnd) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=do),data=gam.dur.2009)
m2010<-gam(log(towhours+1) ~ area.hypoxic +s(jd) + s(depth) + s(ppnd) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=do),data=gam.dur.2010)
m2010<-gam(log(towhours+1) ~ area.hypoxic +s(jd) + s(depth) + s(ppnd) + s(cent_lon,cent_lat) + s(cent_lon,cent_lat, by=do),data=gam.dur.2010)



par(mfrow=c(2,3))
plot(m2004, select=2, shade=T, all.terms=T, scale=0,
     xlab="DO (mg/L)")
plot(m2004, select=3, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
plot(m2004, select=4, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
plot(m2004, select=1, shade=T, all.terms=T, scale=0,
     xlab="Julian Day")

#Get slope coefficients for gom2 model
pred<-predict(m2004, type='terms')
head(pred)
pred.slope<-(pred[,6]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))

par(mfrow=c(1,2))
vis.gam(m2004, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur.2004$cent_lon, gam.dur.2004$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur.2004$cent_lon[pred.slope>=0],gam.dur.2004$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='blue')
symbols(gam.dur.2004$cent_lon[pred.slope<=0],gam.dur.2004$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='red')
map("worldHires", fill=T, col="grey",add=T)

### 2005 ####
#############

par(mfrow=c(2,3))
plot(m2005, select=2, shade=T, all.terms=T, scale=0,
     xlab="DO (mg/L)")
plot(m2005, select=3, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
plot(m2005, select=4, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
plot(m2005, select=1, shade=T, all.terms=T, scale=0,
     xlab="Julian Day")

#Get slope coefficients for gom2 model
pred<-predict(m2005, type='terms')
head(pred)
pred.slope<-(pred[,6]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))

par(mfrow=c(1,2))
vis.gam(m2005, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur.2005$cent_lon, gam.dur.2005$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur.2005$cent_lon[pred.slope>=0],gam.dur.2005$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='blue')
symbols(gam.dur.2005$cent_lon[pred.slope<=0],gam.dur.2005$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='red')
map("worldHires", fill=T, col="grey",add=T)


#########
##2006###
#########

par(mfrow=c(2,3))
plot(m2006, select=2, shade=T, all.terms=T, scale=0,
     xlab="DO (mg/L)")
plot(m2006, select=3, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
plot(m2006, select=4, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
plot(m2006, select=1, shade=T, all.terms=T, scale=0,
     xlab="Julian Day")

#Get slope coefficients for gom2 model
pred<-predict(m2006, type='terms')
head(pred)
pred.slope<-(pred[,6]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))

par(mfrow=c(1,2))
vis.gam(m2006, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur.2006$cent_lon, gam.dur.2006$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur.2006$cent_lon[pred.slope>=0],gam.dur.2006$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='blue')
symbols(gam.dur.2006$cent_lon[pred.slope<=0],gam.dur.2006$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='red')
map("worldHires", fill=T, col="grey",add=T)

#########
##2007###
#########

par(mfrow=c(2,3))
plot(m2007, select=2, shade=T, all.terms=T, scale=0,
     xlab="DO (mg/L)")
plot(m2007, select=3, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
plot(m2007, select=4, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
plot(m2007, select=1, shade=T, all.terms=T, scale=0,
     xlab="Julian Day")

#Get slope coefficients for gom2 model
pred<-predict(m2007, type='terms')
head(pred)
pred.slope<-(pred[,6]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))

par(mfrow=c(1,2))
vis.gam(m2007, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur.2007$cent_lon, gam.dur.2007$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur.2007$cent_lon[pred.slope>=0],gam.dur.2007$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='blue')
symbols(gam.dur.2007$cent_lon[pred.slope<=0],gam.dur.2007$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='red')
map("worldHires", fill=T, col="grey",add=T)

#########
##2008###
#########

par(mfrow=c(2,3))
plot(m2008, select=2, shade=T, all.terms=T, scale=0,
     xlab="DO (mg/L)")
plot(m2008, select=3, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
plot(m2008, select=4, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
plot(m2008, select=1, shade=T, all.terms=T, scale=0,
     xlab="Julian Day")

#Get slope coefficients for gom2 model
pred<-predict(m2008, type='terms')
head(pred)
pred.slope<-(pred[,6]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))

par(mfrow=c(1,2))
vis.gam(m2008, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur.2008$cent_lon, gam.dur.2008$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur.2008$cent_lon[pred.slope>=0],gam.dur.2008$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='blue')
symbols(gam.dur.2008$cent_lon[pred.slope<=0],gam.dur.2008$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='red')
map("worldHires", fill=T, col="grey",add=T)

#########
##2009###
#########

par(mfrow=c(2,3))
plot(m2009, select=2, shade=T, all.terms=T, scale=0,
     xlab="DO (mg/L)")
plot(m2009, select=3, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
plot(m2009, select=4, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
plot(m2009, select=1, shade=T, all.terms=T, scale=0,
     xlab="Julian Day")

#Get slope coefficients for gom2 model
pred<-predict(m2009, type='terms')
head(pred)
pred.slope<-(pred[,6]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))

par(mfrow=c(1,2))
vis.gam(m2009, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur.2009$cent_lon, gam.dur.2009$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur.2009$cent_lon[pred.slope>=0],gam.dur.2009$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='blue')
symbols(gam.dur.2009$cent_lon[pred.slope<=0],gam.dur.2009$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='red')
map("worldHires", fill=T, col="grey",add=T)

#########
##2010###
#########

par(mfrow=c(2,3))
plot(m2010, select=2, shade=T, all.terms=T, scale=0,
     xlab="DO (mg/L)")
plot(m2010, select=3, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
plot(m2010, select=4, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
plot(m2010, select=1, shade=T, all.terms=T, scale=0,
     xlab="Julian Day")

#Get slope coefficients for gom2 model
pred<-predict(m2010, type='terms')
head(pred)
pred.slope<-(pred[,6]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))

par(mfrow=c(1,2))
vis.gam(m2010, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur.2010$cent_lon, gam.dur.2010$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur.2010$cent_lon[pred.slope>=0],gam.dur.2010$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='blue')
symbols(gam.dur.2010$cent_lon[pred.slope<=0],gam.dur.2010$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='red')
map("worldHires", fill=T, col="grey",add=T)

#########
##2011###
#########

par(mfrow=c(2,3))
plot(m2011, select=2, shade=T, all.terms=T, scale=0,
     xlab="DO (mg/L)")
plot(m2011, select=3, shade=T, all.terms=T, scale=0,
     xlab="Depth (m)")
plot(m2011, select=4, shade=T, all.terms=T, scale=0,
     xlab="Price per pound (USD)")
plot(m2011, select=1, shade=T, all.terms=T, scale=0,
     xlab="Julian Day")

#Get slope coefficients for gom2 model
pred<-predict(m2011, type='terms')
head(pred)
pred.slope<-(pred[,6]) #number is the column or specific term you want to produce plots for
hist(pred.slope, main='All Slopes')
max.t<-max(abs(pred.slope))

par(mfrow=c(1,2))
vis.gam(m2011, view=c("cent_lon", "cent_lat"), 
        plot.type="contour", type="response", contour="black",
        color="topo", n.grid=50, too.far=0.2,
        main="", ylab="Latitude", xlab="Longitude")
map("worldHires", fill=T, col="grey",add=T)



plot(gam.dur.2011$cent_lon, gam.dur.2011$cent_lat, pch="", 
     ylab="Latitude", xlab="Longitude")
# vis.gam(la2b, view=c("cent_lon", "cent_lat"), 
#         plot.type="contour", type="link", contour="black",
#         color="gray", n.grid=50, too.far=0.5,
#         main="", ylab="Latitude", xlab="Longitude")
symbols(gam.dur.2011$cent_lon[pred.slope>=0],gam.dur.2011$cent_lat[pred.slope>=0],
        circle=pred.slope[pred.slope>=0],inches=0.2*max(pred.slope[pred.slope>=0])/max.t,add=T,fg='blue')
symbols(gam.dur.2011$cent_lon[pred.slope<=0],gam.dur.2011$cent_lat[pred.slope<=0],
        circle=abs(pred.slope[pred.slope<=0]),inches=0.2*max(abs(pred.slope[pred.slope<=0]))/max.t,add=T,fg='red')
map("worldHires", fill=T, col="grey",add=T)
```


Friday, 20 September 2013
-------------------------

As per JKC I need to make a:
1. Table with number of tows and number of vessels by year (this would be a good table for the paper as well).
2. Maps similar to the first figure of the paper, but that show some random subset of the ELB data (20%?) so you can see the spatial structure a little better
3. Maps of the interpolated area of hypoxia by year (the raw interpolation and the smoothed interpolation) to show how we assigned the DO to tows. 



### Table of tows and vessels for manuscript

```{r}
vms.dat=read.csv("C:/Users/kevin.purcell/Documents/NOAA_SpatialEffort_MS/analysis/vms_analysis/elbdata-20120911.csv", 
                 sep=",",header=T)
vms.20<-subset(vms.dat, vms.dat$towhours<=20)
vms.20<-subset(vms.20, vms.20$mon>=5&vms.20$mon<=8)
```

Year|Vessels|Tows
----|----|-------
2004|`r length(unique(vms.20$boxId[vms.20$yr==2004]))`|`r length(vms.20$towhours[vms.20$yr==2004])`
2005|`r length(unique(vms.20$boxId[vms.20$yr==2005]))`|`r length(vms.20$towhours[vms.20$yr==2005])`
2006|`r length(unique(vms.20$boxId[vms.20$yr==2006]))`|`r length(vms.20$towhours[vms.20$yr==2006])`
2007|`r length(unique(vms.20$boxId[vms.20$yr==2007]))`|`r length(vms.20$towhours[vms.20$yr==2007])`
2008|`r length(unique(vms.20$boxId[vms.20$yr==2008]))`|`r length(vms.20$towhours[vms.20$yr==2008])`
2009|`r length(unique(vms.20$boxId[vms.20$yr==2009]))`|`r length(vms.20$towhours[vms.20$yr==2009])`
2010|`r length(unique(vms.20$boxId[vms.20$yr==2010]))`|`r length(vms.20$towhours[vms.20$yr==2010])`
2011|`r length(unique(vms.20$boxId[vms.20$yr==2011]))`|`r length(vms.20$towhours[vms.20$yr==2011])`


### New map with random subset of data
```{r}
#yearly data sets
vms.2004<-subset(vms.20, vms.20$yr==2004)
vms.2005<-subset(vms.20, vms.20$yr==2005)
vms.2006<-subset(vms.20, vms.20$yr==2006)
vms.2007<-subset(vms.20, vms.20$yr==2007)
vms.2008<-subset(vms.20, vms.20$yr==2008)
vms.2009<-subset(vms.20, vms.20$yr==2009)
vms.2010<-subset(vms.20, vms.20$yr==2010)
vms.2011<-subset(vms.20, vms.20$yr==2011)

#random samples of those years
rand.2004<-vms.2004[sample(1:nrow(vms.2004), 31, replace=F),]
rand.2005<-vms.2005[sample(1:nrow(vms.2005), 121, replace=F),]
rand.2006<-vms.2006[sample(1:nrow(vms.2006), 263, replace=F),]
rand.2007<-vms.2007[sample(1:nrow(vms.2007), 417, replace=F),]
rand.2008<-vms.2008[sample(1:nrow(vms.2008), 672, replace=F),]
rand.2009<-vms.2009[sample(1:nrow(vms.2009), 798, replace=F),]
rand.2010<-vms.2010[sample(1:nrow(vms.2010), 412, replace=F),]
rand.2011<-vms.2011[sample(1:nrow(vms.2011), 596, replace=F),]

#Draw maps
par(mfrow=c(2,4))
plot(rand.2004$slat ~ rand.2004$slon, col="lightskyblue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2004")
map("worldHires", fill=T, col="green",add=T)
plot(rand.2005$slat ~ rand.2005$slon, col="lightskyblue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2005")
map("worldHires", fill=T, col="green",add=T)
plot(rand.2006$slat ~ rand.2006$slon, col="lightskyblue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2006")
map("worldHires", fill=T, col="green",add=T)
plot(rand.2007$slat ~ rand.2007$slon, col="lightskyblue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2007")
map("worldHires", fill=T, col="green",add=T)
plot(rand.2008$slat ~ rand.2008$slon, col="lightskyblue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2008")
map("worldHires", fill=T, col="green",add=T)
plot(rand.2009$slat ~ rand.2009$slon, col="lightskyblue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2009")
map("worldHires", fill=T, col="green",add=T)
plot(rand.2010$slat ~ rand.2010$slon, col="lightskyblue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2010")
map("worldHires", fill=T, col="green",add=T)
plot(rand.2011$slat ~ rand.2011$slon, col="lightskyblue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2011")
map("worldHires", fill=T, col="green",add=T)
```

```{r}

#full sample maps
par(mfrow=c(2,4))
plot(vms.dat$slat[vms.dat$yr==2004] ~ vms.dat$slon[vms.dat$yr==2004], col="blue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2004")
map("worldHires", fill=T, col="green",add=T)
plot(vms.dat$slat[vms.dat$yr==2005] ~ vms.dat$slon[vms.dat$yr==2005], col="blue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2005")
map("worldHires", fill=T, col="green",add=T)
plot(vms.dat$slat[vms.dat$yr==2006] ~ vms.dat$slon[vms.dat$yr==2006], col="blue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2006")
map("worldHires", fill=T, col="green",add=T)
plot(vms.dat$slat[vms.dat$yr==2007] ~ vms.dat$slon[vms.dat$yr==2007], col="blue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2007")
map("worldHires", fill=T, col="green",add=T)
plot(vms.dat$slat[vms.dat$yr==2008] ~ vms.dat$slon[vms.dat$yr==2008], col="blue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2008")
map("worldHires", fill=T, col="green",add=T)
plot(vms.dat$slat[vms.dat$yr==2009] ~ vms.dat$slon[vms.dat$yr==2009], col="blue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2009")
map("worldHires", fill=T, col="green",add=T)
plot(vms.dat$slat[vms.dat$yr==2010] ~ vms.dat$slon[vms.dat$yr==2010], col="blue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2010")
map("worldHires", fill=T, col="green",add=T)
plot(vms.dat$slat[vms.dat$yr==2011] ~ vms.dat$slon[vms.dat$yr==2011], col="blue", pch=1,
     xlim=c(-97.2,-88),
     ylim=c(25.5,30.5),
     xlab="Lon",
     ylab="Lat",
     main="2011")
map("worldHires", fill=T, col="green",add=T)



